Child Signals of Need in Utila, Honduras

### 2024-08-29 Paper.qmd
Added to and edited the methods, results, and discussion sections of the paper.
Cleaned up the analysis script.

We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals in response to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment. Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.

@@

Conflicts of interest within families are largely structured by the likelihood an individual shares their genes with another. As a sexually reproducing species, humans have a 50% chance of passing down a given gene to each child. Each child, however, shares more genes with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%). Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in other siblings, whom parents are expected to value identically, all else being equal [@trivers1974].

@@

## Crying

@@

## Withdrawal

@@

## Temper tantrums and self-harm

@@

## *Study population*

This study was done in Utila, Honduras, one of the country’s three major Bay Islands in the Western Caribbean. Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community made up of expats and their descendants, and 2) Spanish-speaking immigrants from the mainland. Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000]. Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of expat islanders and westerners [@davis2014].

During focus groups held in July 2022, parents indicated high levels of sibling competition and various ways children might signal need. Both Spanish-speaking immigrants and English-speaking island natives reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes. In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home. Spanish-speaking mothers reported that refusing to do homework was a common tactic, while English-speaking mothers emphasized running away and self-harm (e.g., a child slamming their head on the ground) to a greater degree. In response to these hypothesized signals, discipline was common. Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, something that may be partially mediated by differences in signaling among children in addition to information caretakers gained from cues.

## *Sampling*

Participants for this study were recruited by a variety of means. JH recruited early participants through convenience and snowball sampling. Participants were also recruited via Facebook posts, and fliers were sent home with children the Utila's Methodist school.

@@

## *Participants*

All participants for the survey portion of the study met the inclusion criteria. Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in 357 of 375 caregiver/child dyads (our imputation).

@@

## *Ethics*

All adult participants provided informed consent. Surveys were designed to last about hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.

@@

## *Survey*

@@

### *Frequency of child signaling, alloparenting contributions, and conflict with caretaker*

We used 6-point Likert scales to get caregiver reports of 1) child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior. See Table 1 for exact wording.

|                                 |                                                                                                                   |
|-------------------------|-----------------------------------------------|
| Variable                        | Survey question wording                                                                                           |
| Crying Frequency                | How often does your child cry?                                                                                    |
| Sad Frequency                   | How often is your child sad?                                                                                      |
| Tantrum Frequency               | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency          | How often does your child leave the house without telling you when they get mad?                                  |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child?                                                                  |
| Child Alloparenting Frequency   | How often does your child take care of their other siblings?                                                      |

**Table 1**. Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day.

As these responses reflect frequencies of actual behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors. The original wording of the Likert scale and the frequency conversion are shown in Table 2.

|                                                  |                               |
|------------------------------------------------|-----------------------------------------------|
| Likert scale wording                             | Count value (times per month) |
| Never                                            | 0                             |
| Once a month or less                             | 1                             |
| More than once a month but less than once a week | 3                             |
| More than once a week but not daily              | 8                             |
| Daily                                            | 30                            |
| Multiple times per day                           | 60                            |

**Table 2.** Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN.

From these count variables, we created three additional variables to be used, primarily, as our main dependent variables in analyses. SignalFreq was the sum of the frequencies of child sadness, crying, and temper tantrums. SignalFreqMax was the frequency of the most common of these three signals. SignalCost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., SignalCost = SadFreqN + CryFreqN\*2 + TantrumFreqN\*3). The frequency of running away was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.

### *Relative neediness of children and relative investment in children within the family*

We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table 3 for exact wording.

|                     |                                                                                                  |
|-------------------------|-----------------------------------------------|
| Variable            | Survey question wording                                                                          |
| Relative Need       | How much time, effort, and resources does this child need compared to your other children?       |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

**Table 3.** Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more.

### *Last instance of child signaling*

We also asked caretakers about the last time each participating child was sad or cried. This involved providing parents with a free response text box to describe the cause of the emotion that lead to the display. We then asked how the parent responded to their child. The researcher giving the survey would then bin responses into one of 31 categories which included positive, negative, and neutral responses. For the full list of options see Table S1.

### *Child anthropometrics*

@@

We computed each child's height, weight, and BMI Zscores with respect to the WHO child growth standards (cite).

### *Perceptions of child immune function (or current susceptibility to illness)*

Caretakers were also asked three questions using a five point Likert scale about their child's current health which involved the likelihood of them becoming sick. We then inverted the scores for two measures so that a score of 5 corresponded with the lowest susceptibility to illness before averaging the scores of these to create the CurrentHealthMean variable. Means were still calculated if 1-2 values had an NA (only xxx cases), with the numerator being the number of variables with data. See Table 4 for the wording of these variables.

|                |                                                                                       |
|-------------------------|-----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

: **Table 4.** Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness.

### *Demographic questions*

@@

CHANGE TO TOTAL CHILDCARE

DECIDE on alloparentingFreqN2

@@

## *Statistical analyses*

We fit generalized linear mixed models for exploratory analyses with signal frequency and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with householdID as a random effect.

@@

## *Data availability*

@@

```{r sumtab_children}
#| echo: false
#| label: sumtab_children
#| fig-cap: "Summary statistics for child study variables."

vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

kable(custom.summarize(d2, vars_c)) %>% kable_styling()

@@

```{r sumtab_adults}
#| echo: false
#| label: sumtab_cadults
#| fig-cap: "Summary statistics for adult and household study variables."

@@

kable(custom.summarize(d2, vars_c)) %>% kable_styling()

@@

>>>>

We argue that older children (i.e., those older than infants) have evolved strategies to tailor their signals to 1) their valuation of a given resource, 2) the extent of conflicts of interest with caretakers, and 3) the presence of existing information about a child's need or potential to benefit from a resource so that they can credibly convey they are worth additional investment. Such strategies can lead to some combination of verbal requests for more support, facial sadness, crying, withdrawal, temper tantrums and self-harm, all of which may be effective in certain circumstances and are sometimes successful in eliciting more support from caretakers, despite a child's low status and formidability.

@@

Conflicts of interest within families are largely structured by the likelihood an individual shares their genes with another. As a sexually reproducing species, humans have a 50% chance of passing down a given gene to each child. Each child, however, shares more genes with themselves (100%) than with their full siblings (50% on average), half-siblings (25% on average), and stepsiblings (0%). Due to this asymmetry, children are often expected to attempt to elicit more investment from parents than is optimal from the perspective of the parent, while parents are expected to attempt to invest less than is optimal for each child in order to invest in their other children, whom parents are expected to value identically, all else being equal [@trivers1974].

@@

### Crying

@@

### Withdrawal

@@

### Temper tantrums and self-harm

@@

## Study population

This study was done in Utila, Honduras, one of the country’s three major Bay Islands in the Western Caribbean. Participants in the study largely fit into two groups who face different levels of economic and social adversity, 1) a long-standing English-speaking community primarily made up of the descendants of expats and 2) a relatively new community of Spanish-speaking immigrants from the mainland. Compared to those born on Utila, immigrants from the mainland [@davis2014] are more likely to live in crowded and unsanitary conditions, face active discrimination [@garcia2017], and experience economic uncertainty due the predominance of low paying, seasonal jobs primarily related to the island’s diving tourism industry [@garcia2018; @stonich2000]. Such inequality appears to be increasing in response to recent growth in the tourism industry, with the majority of wealth from tourism being enjoyed by a small portion of expat islanders and westerners [@davis2014].

During focus groups held in July 2022, parents indicated high levels of sibling competition and various ways children might signal need. Both Spanish-speaking and English-speaking moms reported that fighting between siblings was extremely common, with fights over toys, gifts, and phone time being the most reported causes. In both populations, crying and various forms of withdrawal were listed as the most common responses to a child not receiving investment or support, followed by temper tantrums and making messes in their home. Spanish-speaking mothers reported that refusing to do homework was a common tactic, while English-speaking mothers emphasized running away and self-harm (e.g., a child slamming their head on the ground) to a greater degree. In response to these hypothesized signals, discipline was common. Although the mothers generally reported that children were treated equally, needier children were described as getting more resources from parents, something that may be partially mediated by differences in signaling among children in addition to information caretakers gained from cues.

## Sampling

Participants for this study were recruited by a variety of means. JH recruited early participants through convenience and snowball sampling. Participants were also recruited via Facebook posts and fliers which were sent home with children the Utila's Methodist school.

@@

## Participants

All participants for the survey portion of the study met the inclusion criteria. Most participants were mothers of at least one biological child in the study, with the participant being the biological mother of the child in xxx out of xxxx caregiver/child dyads.

@@

## Ethics

All adult participants provided informed consent. Surveys were designed to last about an hour plus 20 minutes for each additional child. Adults were paid \$XXXXX for the survey portion of the study.

@@

## Survey

@@

### Frequency of child signaling, alloparenting contributions, and conflict with caretaker

We used 6-point Likert scales to get caregiver reports on the frequencies of 1) various types of child signaling, 2) interpersonal conflict with each participating child, 3) and child alloparenting behavior. See @tbl-freqvars for exact wording. Caretakers were often not asked about child alloparenting behavior if a child was an only child. For this reason, we gave children an alloparenting frequency of never if they were the only child within a household.

|                                 |                                                                                                                   |
|-------------------------|----------------------------------------------|
| Variable                        | Survey question wording                                                                                           |
| Crying Frequency                | How often does your child cry?                                                                                    |
| Sad Frequency                   | How often is your child sad?                                                                                      |
| Tantrum Frequency               | How often does your child throw temper tantrums? (e.g., scream, throws things, or throws themselves on the floor) |
| Running Away Frequency          | How often does your child leave the house without telling you when they get mad?                                  |
| Parent/Child Conflict Frequency | How often do you have conflicts with your child?                                                                  |
| Child Alloparenting Frequency   | How often does your child take care of their other siblings?                                                      |

: Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day. {#tbl-freqvars}

As these responses reflect frequencies of behavior, we converted these to counts which reflected the number of times per month a child engaged in these behaviors for analyses. The original wording of the Likert scale and the frequency conversion are shown in @tbl-convertedfreqs.

|                                                  |                               |
|------------------------------------|------------------------------------|
| Likert scale wording                             | Count value (times per month) |
| Never                                            | 0                             |
| Once a month or less                             | 1                             |
| More than once a month but less than once a week | 3                             |
| More than once a week but not daily              | 8                             |
| Daily                                            | 30                            |
| Multiple times per day                           | 60                            |

: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN. {#tbl-convertedfreqs}

From these count variables, we created three additional signaling variables to be used, primarily, as our main dependent variables in analyses. The frequency of signaling was the sum of the frequencies of child sadness, crying, and temper tantrums. Maximum signal frequency was the frequency of the most common of these three signals. Signal cost was the sum of the frequencies of these three signals weighted by their relative costliness (i.e., Signal cost = frequency of sadness + frequency of crying * 2 + frequency of temper tantrums * 3). Although we collected data on the frequency of running away, it was not included in the creation of any of these variables due to uncertainty over whether it was likely to be a credible signal due to the relative safety of Utila compared to other areas in which running away has been found to be a successful strategy for children.

### Relative neediness of children and relative investment in children within the family

We also used 6-point Likert scales to get parental reports of 1) the relative neediness of each participating child and 2) the relative investment they provide to each participating child compared to the rest of their children. See Table @tbl-relvars for exact wording.

|                     |                                                                                                  |
|-------------------------|----------------------------------------------|
| Variable            | Survey question wording                                                                          |
| Relative Need       | How much time, effort, and resources does this child need compared to your other children?       |
| Relative Investment | How much time, effort, and resources do you put into this child compared to your other children? |

: Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. {#tbl-relvars}

### Causes and responses to the last instance of child signaling

We also asked caretakers about the last time each participating child was sad or cried. This involved asking what the caretaker thought caused the emotion or display and recording their answer in a text box. We then asked how the parent responded to their child. The researcher giving the survey would then bin responses into at least one of 31 categories which included positive, negative, and neutral responses (from the child's perspective). For the full list of options see Table S1.

From this data, we created a parental response variable, which was coded as 1 for a positve response, 0 for a neutral response, and -1 for a negative response (all from the child's perspective). MG also coded each cause for the presence or absence of 17 various elements. Of these 4 were use in analyses: whether or not the cause involved 1) the child being asked to do an unwanted task, 2) punishment, 3) conflict within the family, and 4) discomfort, pain, injury, or illness.

### Child anthropometrics

@@

We computed each child's height, weight, and BMI Zscores with respect to the CDC child growth standards (cite).

### Perceptions of child immune function (or current susceptibility to illness)

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable. In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean a numerator of 2. See Table @tbl-healthvars for the wording of these variables.

|                |                                                                                       |
|-------------------------|----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness. {#tbl-healthvars}

### Demographic questions

@@

OLD AND CURRENTLY NOT USING

@@

## Statistical analyses

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables invovled over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with householdID as a random effect.

@@

## Data availability

@@

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures"
)

kable(custom.summarize(modeldf, vars_c)) %>% kable_styling()

@@

# CONFIRM: should this be household level?
```{r sumtab_adults}
#| label: tbl-sumtab_adults
#| tbl-cap: Summary statistics for adult and household study variables.

@@

kable(custom.summarize(modeldf, vars_c)) %>% kable_styling()

@@

## Distributions of signaling behavior, alloparenting effort, and conflict

Child sadness, crying, and temper tantrums were reasonably common. Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) of children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\])

Conflict between children and caretakers was also common, with `r sum(na.omit(modeldf$ConflictFreqN >= 8))` children (`r signif(sum((na.omit(modeldf$ConflictFreqN >= 8)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%) having conflicts with parents more than once a week. As with the frequencies of various types of signaling, many children (`r sum(na.omit(modeldf$ConflictFreqN == 0))` \[`r signif(sum((na.omit(modeldf$ConflictFreqN == 0)) / sum(!is.na(modeldf$ConflictFreqN)) * 100), 2)`%\]) were described as never having conflicts with parents (stat).

The frequency of alloparenting effort was bimodally distributed, with `r sum(na.omit(modeldf$AlloparentingFreqN == 0))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN == 0)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to never engage in it (`r sum(na.omit(modeldf$NumberOfChildren==1))` instances were due to a child being an only child) and `r sum(na.omit(modeldf$AlloparentingFreqN >= 30))` children (`r signif(sum((na.omit(modeldf$AlloparentingFreqN >= 30)) / sum(!is.na(modeldf$AlloparentingFreqN)) * 100), 2)`%) being reported to help with childcare at least once a day.

::: {#fig-barplot_SignalFreq}
```{r}
barplot_SignalFreq
```

Stacked bar plot showing the frequencies of child sadness, crying, and temper tantrums per month as reported by primary caretaker. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day." 
:::

::: {#fig-signal_alluvial_plot}
```{r}
signal_alluvial_plot
```

Alluvial plot showing the combinations of signaling frequencies. Alluvium width equals the number of individiuals with the same frequencies for the three signals/states. Stratum height equals the count of individuals at a given frequency. Frequency key: Low includes "never" and "once a month or less," mediums includes "more than once a month but less than once a week" and "more than once a week but not daily," high includes "daily," and  "multiple times per day."
:::

::: {#fig-barplot_ConflictFreq}
```{r}
barplot_conflict
```

Stacked bar plot showing the frequency of conflict between child and primary caregiver as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day." 
:::

::: {#fig-barplot_AlloparentingFreq}
```{r}
barplot_alloparenting
```

Stacked bar plot showing the frequency of conflict a child's alloparenting effort as reported by primary caretaker. Color corresponds to sex of the child. Frequency key: 0 corresponds to "never," <=1 corresponds to "once a month or less," 2-3 corresponds to "more than once a month but less than once a week," 4-29 corresponds to "more than once a week but not daily," 30 corresponds to "daily," and >30 corresponds to "multiple times per day."
:::

## Signaling behavior is negatively associated with age

For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.

Age had largest effect on crying (`r stats$mcryH$ChildAge$str`), with the model predicting crying frequency dropping to half of the age 5 value by around age 8 (visual inspection). The effect of age on temper tantrums (`r stats$mtantrumH$ChildAge$str`) was smaller with a more linear slope.

```{r}
#| label: tbl-main_models
#| tbl-cap: Regression tables for our main exploratory models.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare"
)
modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

::: {#fig-effect_age}
```{r}
signaling_data_plot_age
```

The effect of age on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

## Signaling behavior is positively associated with conflict with primary caretakers

There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.

# FIX: y axis extends farther for tantrum model (reducing this manually breaks error shading)

::: {#fig-effect_conflict}
```{r}
signaling_data_plot_conflict
```

The effect of the frequency of conflict with one's primary caretaker on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

## Interaction between child sex and child alloparenting

::: {#fig-effect_allo_by_sex}
```{r}
signaling_data_plot_allo_by_sex
```

The effect of an alloparenting effort by sex interaction on specific signal frequencies, summed signal frequency, the frequency of a child's most common signal, and the summed cost of signaling behavior. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-main_models for the corresponding regression table.
:::

For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function. The slope of these relationships were curvilinear and positive for girls and flat or slightly negative for boys. In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys. See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.

## Summed signal cost, crying frequency, and tantrum frequency are negatively associated with body fat

There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots. Subscapular and tricep skinfold thicknesses were also significant predictors of summed signal cost and tantrum frequency when included in the same models as the sole measure of body fat. See xxxxSI TABLESxxxx for regression tables.

Height Z-scores, weight Z-scores, BMI Z-scores, and body fat percentage residuals were not significant predictors of signaling behavior in models with the same control variables. See xxxxSI TABLESxxxx for regression tables.

```{r}
#| label: tbl-bodyfat_models
#| tbl-cap: Regression tables for the models examining the effect of body fat on signaling.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)"
)
modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

::: {#fig-effect_bodyfat}
```{r}
signaling_data_plot_bodyfat
```

The effect of body fat (scaled composite measure of subscapular and tricep skinfold thicknesses) on A) the summed cost of signaling behavior, B) the frequency of crying, and C) the frequency of temper tantrums. Shaded areas are 95% CIs (only taking the uncertainty of fixed-effect parameters into account). Each dot is 1 child. Overlapping points are represented by the transparency of the point (lower transparency equals more values at a given coordinate). See @tbl-bodyfat_models for the corresponding regression table.
:::

## Positive associations between child age and signaling frequency and child neediness

Due to the rarity of primary caretakers describing a child as "much less" (`r sum(na.omit(modeldf$RelativeNeed == "Much less"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much less"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) or "much more" (`r sum(na.omit(modeldf$RelativeNeed == "Much more"))` \[`r signif((sum(na.omit(modeldf$RelativeNeed == "Much more"))/ sum(!is.na(modeldf$RelativeNeed)) * 100), 2)`%\]) needy compared to their other children, we collapsed "less" and "much less" into one "less" factor and "more" and "much more" into one "more" factor. The end result was `r rel_need_table[1]` less needy children, `r rel_need_table[2]` children who were as needy as their siblings, and `r rel_need_table[3]` more needy children.

There were significant positive relationships between perceived relative neediness within the family and both the summed frequency of child signaling and child age when controlling for child sex, household demographics, the frequency of conflict between child and caretaker, and the frequency of a child providing allocare. There was also a positive relationship between summed signal cost and perceived relative neediness within the family when controlling for the same variables; however, this association was not significant. See @tbl-need_models for the corresponding regression table.

#FIX: how to use coeftest with this
```{r}
#| label: tbl-need_models
#| tbl-cap: Regression tables for the models examining relative perceptions of child need within the family as a dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)"
)
modelsummary(list(SignalFreq = mN1, SignalCost = mN2), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

::: {#fig-effect_relativeneed}
```{r, fig.width= 10}
signaling_plot_need
```

Plots showing the likelihood of child being more needy, as needy, or less needy than a caretaker's other children. A: Effect of child age on relative neediness B: Effect of summed signaling frequency on relative neediness.
:::

## Parental responses to last signal of need

There were `r response_overlap[2,2]` instances in which a caregiver's response involved acts which were both negative and positive from the child's perspective. These were excluded from analyses due to their rarity.

There were no significant relationships between, child age, child sex, the number of children in the household, and log family income and the valence of parental reponses (positive, neutral, or negative) from the child's perspective. However, primary caretakers were significantly more likely to respond positively towards a child when they attributed the cause of signaling to discomfort, pain, injury, or illness (stat), and significantly less likely to respond positively when they attributed the cause of signaling to child punishment (stat). See @fig-effect_last_signal_response and @tbl-last_signal_model.
